---
title: "Mark Dunning"
author: "Revised Analysis of the Pasilla dataset"
date: "29th July 2015"
output: pdf_document
---

The counts for the pasilla dataset were read from the pasilla data package.

```{r dataload,echo=FALSE,message=FALSE}
library(edgeR)
datafile = system.file( "extdata/pasilla_gene_counts.tsv", package="pasilla" )
```

The first few lines of the file are shown.


```{r readCounts}
pasillaCountTable = read.table( datafile, header=TRUE, row.names=1 )

head(pasillaCountTable)
```

A design matrix was used to compare treated and untreated samples.

```{r makeDesign}
pasillaDesign = data.frame(
  row.names = colnames( pasillaCountTable ),
  condition = c( "untreated", "untreated", "untreated",
                 "untreated", "treated", "treated", "treated" ),
  libType = c( "single-end", "single-end", "paired-end",
               "paired-end", "single-end", "paired-end", "paired-end" ) )


pasillaDesign
pairedSamples = pasillaDesign$libType == "paired-end"
countTable = pasillaCountTable[ , pairedSamples ]
condition = pasillaDesign$condition[ pairedSamples ]
```

```{r}
y <- DGEList(counts=countTable,group=condition)
```
Prior to filtering, there are `r nrow(y)` genes


```{r}
countsPerMillion <- cpm(y)
summary(countsPerMillion)
#'summary' is a useful function for exploring numeric data; eg. summary(1:100)
countCheck <- countsPerMillion > 1
head(countCheck)
keep <- which(rowSums(countCheck) >= 2)
y <- y[keep,]
summary(cpm(y))
```

After filtering, `r nrow(y)` genes remain.

The analysis will use `r table(pasillaDesign$condition)[1]`  Treated and `r table(pasillaDesign$condition)[2]`  Untreated samples. Normalisation was performed with a standard edgeR protocol

```{r normalise}

y <- calcNormFactors(y)
y <- estimateCommonDisp(y)
y <- estimateTagwiseDisp(y)
```


Differential expression was performed between treated and untreated samples using the exact test in edgeR.
The top hits are shown below.
```{r doDE}
et <- exactTest(y)
topTags(et)
p <- 0.01
summary(de <- decideTestsDGE(et, p=p))
detags <- rownames(y)[as.logical(de)]
```

The total number of differentially-expressed genes at a cutoff of `r p` was `r sum(abs(de)==1)`, and `r sum(de==1)` genes were up-regulated. The
logFC and CPM of these differentially-expressed genes is shown below.


```{r plotSmear}
plotSmear(et, de.tags=detags)
abline(h = c(-2, 2), col = "blue")
```

